First steps in synthetic algebraic geometry

### 2017-03-19 snippets/grassmannian.tex
More on the Grassmannian

@@

\begin{cor}Die Grassmannsche~$\Gr(V,r)$ ist ein Schema von endlichem
Typ.\end{cor}

\begin{proof}Die Grassmannsche~$\Gr(V,r)$ besitzt eine offene
Überdeckung durch die affinen Schemata~$G_W$ und ist daher ein Schema.

Wenn wir einen Isomorphismus~$V \cong k^n$ wählen, sehen wir, dass
schon~$\binom{n}{r}$ viele dieser affinen Schemata genügen: nämlich diese,
wo~$W$ einer der Standarduntermoduln von~$k^n$ ist (erzeugt durch
Einheitsvektoren).

Denn ist~$U \in \Gr(k^n,r)$, so bildet die Surjektion~$V \to V/U$ die Basis von
mindestens einem dieser Standarduntermoduln auf eine Basis ab und ist daher
bijektiv. (Aus einer surjektiven~$(r \times n)$-Matrix über einem lokalen Ring
kann man stets~$r$ Spalten auswählen, die eine linear unabhängige Familie
bilden.)
\end{proof}

>>>>

\begin{cor}The Grassmannian~$\Gr(V,r)$ is a finitely presented scheme.\end{cor}

\begin{proof}We need to verify that~$\Gr(V,r)$ admits a finite covering by
spectra of finitely presented~$\affl$-algebras. We already know that~$\Gr(V,r)$
can be covered by the open subsets~$G_W$ and that these sets are spectra of
finitely presented algebras. Therefore it remains to prove that finitely many of
these subsets suffice to cover~$\Gr(V,r)$.

In fact, if we choose an isomorphism~$V \cong (\affl)^n$, we see
that~$\binom{n}{r}$ of these subsets suffice: namely those where~$W$ is one of
the standard submodules of~$(\affl)^n$ (generated by the standard basis
vectors). For if~$U \in \Gr((\affl)^n,r)$, the surjection~$V \to V/U$ maps
the basis of at least one of these standard submodules to a basis and is
therefore bijective. This is because from any surjective~$(r \times n)$-matrix
over a local ring one can select~$r$ columns which form an linearly independent
family.
\end{proof}

@@
